---
title: "Week 1 Assignment"
author: "Ben Arancibia"
date: "June 11, 2015"
output: pdf_document
---


####HA 2.1

For each of the following series (from the fma package), make a graph of the data. If transforming seems appropriate, do so and describe the effect.

```{r}
library(fma)
```

a) Monthly total of people on unemployed benefits in Australia (January 1956–July 1992).
```{r}
lambda.benefits <- BoxCox.lambda(dole)
plot(BoxCox(dole, lambda.benefits), main="Monthly People on Benefits", xlab="1956-1992", ylab="Number of people(Box-Cox transformed)")
title(main=paste("Box Cox lambda = ", signif(lambda.benefits, digits=3)), font.main=8, line=-1)
```

b) Monthly total of accidental deaths in the United States (January 1973–December 1978).
```{r}
lambda.deaths <- BoxCox.lambda(usdeaths)
plot(BoxCox(usdeaths, lambda.deaths), main="Monthly Accidental Deaths", xlab="1973-1978", ylab="Number of people(Box-Cox transformed)")
title(main=paste("Box Cox lambda = ", signif(lambda.deaths, digits=3)), font.main=8, line=-1)
```

c) Quarterly production of bricks (in millions of units) at Portland, Australia (March 1956–September 1994).
```{r}
lambda.bricks <- BoxCox.lambda(bricksq)
plot(BoxCox(bricksq, lambda.bricks), main="Quarterly Production Bricks", xlab="1973-1978", ylab="Number of bricks(Box-Cox transformed)")
title(main=paste("Box Cox lambda = ", signif(lambda.bricks, digits=3)), font.main=8, line=-1)
```

####HA 2.3
Consider the daily closing IBM stock prices (data set ibmclose).
a)Produce some plots of the data in order to become familiar with it.
```{r}
plot(ibmclose)
tsdisplay(ibmclose)
```
b)Split the data into a training set of 300 observations and a test set of 69 observations.

c)Try various benchmark methods to forecast the training set and compare the results on the test set. Which method did best?

