---
title: "Week 2 Homework"
author: "Ben Arancibia"
date: "June 20, 2015"
output: pdf_document
---

#### HA 4.1

Electricity consumption was recorded for a small town on 12 randomly chosen days. The following maximum temperatures (degrees Celsius) and consumption (megawatt-hours) were recorded for each day. 

```{r, comment=FALSE,warning=FALSE,message=FALSE}
library(fpp)
```
```{r}
knitr::kable(econsumption)
```

a) Plot the data and find the regression model for Mwh with temperature as an explanatory variable. Why is there a negative relationship?

```{r}
plot(Mwh ~ temp, data=econsumption)
fit  <- lm(Mwh ~ temp, data=econsumption)
summary(fit)
```

It seems that there is a negative relationship because a simple linear model is not appropriate. A non-linear model will be necessary for the data.

b) Produce a residual plot. Is the model adequate? Are there any outliers or influential observations?

```{r}
plot(residuals(fit)~temp, data=econsumption)
```

The model appears to be adequate, there is no apparent pattern. The only thing that might cause an issues is 1 outlier at temperature 23.4.


c) Use the model to predict the electricity consumption that you would expect for a day with maximum temperature 10 and a day with maximum temperature 35. Do you believe these predictions?

```{r}
fcast <- forecast(fit, newdata=data.frame(temp=c(10,35)))
plot(fcast)
```

Based on the predictive intervals and the nearby data points, the predictions are believable.

d) Give prediction intervals for your forecasts. The following R code will get you started:

```{r}
summary(fit)
forecast(fit, newdata=data.frame(temp=c(60)))
```

####HA 5.2
