---
title: "Week 3 Homework"
author: "Ben Arancibia"
date: "June 23, 2015"
output: pdf_document
---

####KJ 7.2

Friedman (1991) introduced several benchmark datasets create by simulation. One of these simulations used the following non--linear equation to create data: 

$y = 10 \sin(\pi x_1x_2) + 20 (x_3 - 0.5)^2 + 10 x_4 + 5 x_5 + N(0, \sigma^2)$

```{r}
library(mlbench)
library(caret)
set.seed(200)
trainingData <- mlbench.friedman1(200, sd = 1)
trainingData$x <- data.frame(trainingData$x)
featurePlot(trainingData$x, trainingData$y)
testData <- mlbench.friedman1(5000, sd = 1)
testData$x <- data.frame(testData$x)
```

Tune several models on these data. For example:

```{r}
set.seed(921)
knnModel <- train(x = trainingData$x, y = trainingData$y, method = "knn",
                  preProc = c("center", "scale"),
                  tuneLength = 10)
knnModel
knnPred <- predict(knnModel, newdata = testData$x)
postResample(pred = knnPred, obs = testData$y)
```

Which models appear to give the best performance? Does MARS select the informative predictors (those named X1–X5)?

K-nearest neighbors models are better when predictors and the response relies on the samples' proximity in the predictor space.