---
title: "Week 5 Homework"
author: "Ben Arancibia"
date: "July 8, 2015"
output: pdf_document
---

```{r}
library(fpp)
```

####HA 6.1

1) Show that a 3×5 MA is equivalent to a 7-term weighted moving average with weights of 0.067, 0.133, 0.200, 0.200, 0.200, 0.133, and 0.067.

First do the 7 term weighted moving average
```{r}
weights <- (c(0.067, 0.133, 0.200, 0.200, 0.200, 0.133, 0.067))

ma7 <- ma(weights, order=7)

ma7
```

Next do the 3x5 MA
```{r}
ma5 <- ma(weights, order=5)
ma3x5 <- ma(ma5, order=3)
```

The reason that 3x5 is equivalent to the 7 term weight moving average is because combinations of moving averages result in weighted moving averages. 

####HA 6.2
The data below represent the monthly sales (in thousands) of product A for a plastics manufacturer for years 1 through 5 (data set plastics).
```{r}
plastics
```

a) Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend?

```{r}
plot(plastics)
```

There is seasonal and upward trend for the product A.

b) Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

```{r}
decomp <- decompose(plastics, type="multiplicative")

plot(decomp)
```

c) Do the results support the graphical interpretation from part (a)?

Yes the results show there is an unchanged seasonanility (third graphic), but there is a tip in the trend that was not noticed in the first plot.

d) Compute and plot the seasonally adjusted data.

```{r}
sa <- seasadj(decomp)
plot(sa)
```


e) Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

```{r}
x <- plastics
x[18] <- x[18] + 500
sa2 <- seasadj(stl(x, s.window="periodic"))
sa3 <- seasadj(stl(x, s.window="periodic", robust=TRUE))
plot(sa, col="blue", ylim=range(sa,sa2,sa3))
lines(sa2,col="red")
lines(sa3, col="green")
```

If robust=FALSE, the seasonal adjusted series changes across the whole data set. Notice that there are a lot more differences in the red line compared to the blue line. If robust=TRUE, only the outlying point changes noticeably. There are not that many differences from the blue and green lines compared to the blue line.

f) Does it make any difference if the outlier is near the end rather than in the middle of the time series?

```{r}
#Repeat above function

x <- plastics
x[58] <- x[58] + 500 
sa2 <- seasadj(stl(x, s.window="periodic"))
sa3 <- seasadj(stl(x, s.window="periodic", robust=TRUE))
plot(sa, col="blue", ylim=range(sa,sa2,sa3))
lines(sa2,col="red")
lines(sa3, col="green")
```

Having an outlier near the end gives it more impact.

g) Use a random walk with drift to produce forecasts of the seasonally adjusted data.

```{r}
fc <- stlf(plastics, method="naive")
plot(fc)
```

h) Reseasonalize the results to give forecasts on the original scale.

```{r}
fit <- stl(plastics, s.window="periodic", robust=TRUE)
fitadj <- seasadj(fit)
plot(naive(fitadj), xlab="Plastic Product A",
  main="Naive forecasts of seasonally adjusted data")
```


